The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-05-30 paper/ms.tex
discussion

%% One possible explanation for the emergence of the single peak in the
%% homogeneous systems is that some photons close to the surface
%% (a sort of photosphere) can escape with a low number of scatterings,
%% allowing them to stay close to the line's center. Increasing the
%% rotational velocity $V_{\rm max}$ reduces the optical depth making the
%% photosphere region effectively larger, increasing the number of
%% photons escaping close to the line's center.

%% However, for the central emission the transition to a single peak is
%% harder to reach in the range of explored parameters. 
%% In this case the average number of scatterings to escape
%% should remain close to constant as the rotational velocity
%% increases. The rise in intensity at the line's center could instead
%% mean that the scattering in a medium with bulk motion are inefficient
%% in driving photons outside that frequency region.

%% In order to explore our interpretation for these two scenarios we now 
%% quantify the effect of rotation on the number of scatterings.

@@

Gas bulk rotation has a noticeable effect on the morphology of the
 \ly line. In this section we compare our findings to other
 computational results and discuss the implications of these finding
 in the interpretation of observational data.

\cite{Verhamme12} studied \ly using two high resolution simulations
of individual galaxies. The main purpose of the study was to asses the
impact of two different InterStellar Medium (ISM) prescriptions. However, each simulated
galaxy had a disc structure with a clear rotation pattern in the ISM
and inflowing gas int the circum-galactic region. The configuration
had an axial symmetry and they reported a strong dependence of the
escape fraction and the total line intensity as a function of the
$\theta$ angle. In our results, none of these two quantities has a
dependence either on the inclination angle or the rotational
velocity. Therefore, the effect reported by \cite{Verhamme12} is
consistent with being a consequence of the different hydrogen optical
depth for different viewing angles and not as an effect of the bulk
rotation.

The presence of single peaked profiles has been associated to
inflow/outflow dynamics \citep{Verhamme06,DijkstraKramer}. In this
paper we show that gas bulk rotation can also be considered as a
probable origin for that behavior, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserves to be considered in
the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. This highlights
that in order to interpret the observations of \ly emitting galaxies
it is necessary to consider the possible effect of different
inclination angles along the line of sight.

Another important issue is the \ly flux, in several systems correlation
between the \ly flux and properties such as the peak shift has been found
 (cites), in our study we don't find this correlation mainly because the flux remain 
 constant for all the rotational velocities and because the flux of the line
 is independent of the viewing angle. This last feature is shown in 
 Fig.\ref{fig:muhisto} when clearly is seen that the flux of the line
 does not change with the viewing angle $\theta$ as it does in systems with
 anisotropic \ly emission \citep{Zheng2013} and most recently in
  Outflow-Driven Cavities \citep{Behrens2014}.

>>>>

@@

We now compare our findings to other computational results and discuss
its possible implications  for the interpretation of observational data.

{\bf Galaxy simulations with gas rotation}. \cite{Verhamme12} studied \ly
line emission in two high resolution simulations of individual
galaxies. The main purpose of their study was to asses the impact of
two different Inter-Stellar Medium (ISM) prescriptions. However, each
simulated galaxy had a disc structure with a clear rotation pattern in
the ISM and inflowing gas from the circum-galactic region. The configuration
had an axial symmetry and they reported a strong dependence of both
the escape fraction and the total line intensity as a function of the
$\theta$ angle. In our results, none of these two quantities has a
dependence either on the inclination angle or the rotational
velocity. We suggest that he effect reported by \cite{Verhamme12} is
consistent with being a consequence of the different hydrogen optical
depth for different viewing angles and not as an effect of the bulk
rotation.

{\bf Single peaked lines}. The presence of single peaked profiles has
been associated to inflow/outflow dynamics
\citep{Verhamme06,DijkstraKramer}. In this paper we show that gas bulk
rotation can also be considered as a probable origin for that
behavior, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserves to be
considered in the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. This highlights
that in order to interpret the observations of \ly emitting galaxies
it is necessary to consider the possible effect of different
inclination angles along the line of sight.

{\bf Systemic velocities}. There are observational measurements for the
velocity shift between the \ly and other emission lines. In our study
we find that the position of the peak maxima is insensitive to the
rotational velocity of the gas. In our model the only parameter that
can influence the peak maxima position is the optical depth.

{\bf Anisotropic emission}. In our studies the integrated flux is
constant is independent of the viewing angle. This rules out the
possibility of having an observational indication of gas bulk rotation
from measurements of the line equivalent width. A possible way to
obtain an anisotropic flux with the same gas configuration is to allow
for anisotropic emission \citep{Zheng2013,Behrens2014}. We explore
these scenarios in future work (Remolina-Gutierrez et al., in prep.).

{\bf Additional uncertainty on the \ly escape fraction}. Study of
high redshift LAEs in numerical simulation often requires the
estimation of the \ly escape fraction in order to compare their
results against observations
\citep{CLARA,Dayal2012,Forero12,Orsi12,Garel2012}. Most of these
models estimate the escape fraction from the column density of dust and
neutral Hydrogen. The results of our simulation suggest that the
rotational velocity has also to be accounted for, having an impact on
the values of the escape fractions by factors of $\sim 2$.
